---
title: "Homework-6: Pairwise Association"
format: pdf
---

## Problem 1

The New York Times director is investigating the relationship between where a subscriber lives and which sections of the newspaper they read first. Use the collected sample information to answer the following questions:

|          | Politics  | Editorial | Sports |
|:--------:|:---------:|:---------:|:------:|
| City     | 170       | 124       | 90     |
| Suburb   | 120       | 130       | 100    |
| Rural    | 130       | 90        | 88     |

a. At the 95% confidence level, is it reasonable to conclude that the section read first and the community where a reader resides are dependent?

```{r Question 1 A}

politics <- c(170,120,130)
editorial <- c(124,130,90)
sports <- c(90,100,88)
NewYorkTimeData <- data.frame(politics, editorial, sports)

result <- chisq.test(NewYorkTimeData)
p_value <- result$p.value
print(p_value)
alpha <- 0.05

```
> Since the p-value of 0.03242511 is smaller than the alpha value of 0.05, we reject the null hypothesis. We can not be 95% sure that the section read first and the community where a reader resides are dependent.

b. What is the largest chi-square value? What is the meaning of this value?

```{r Question 1 B}

total <- sum(NewYorkTimeData)
row_totals <- rowSums(NewYorkTimeData)
col_totals <- colSums(NewYorkTimeData)
expected <- outer(row_totals, col_totals) / total

# Calculate chi-squared contributions for each cell
chi_squared_contributions <- (NewYorkTimeData - expected)^2 / expected

# Total chi-square statistic
total_chi_square <- sum(chi_squared_contributions)

# Print chi-squared contributions for each cell
print(chi_squared_contributions)
```